---
title: "R Notebook"
output: 04_visium_features
editor_options: 
  chunk_output_type: console
---


Let's load our already qc-d SPE.

```{r}
library(SpatialExperiment)
library(ggspavis)
library(nnSVG)
library(scran)

spe <- readRDS(file.path('visium', 'visium_spe_qc.rds'))
```

# Feature selection

My plan was to use nnSVG to define spatially-variable genes.
Unfortunately, this particular method is quite slow and memory intensive,
So for time, we're going to use a method that doesn't make use of the spatial data.
This will be more similar to the simple variance analysis we previously performed.


```{r}
dec <- modelGeneVar(spe)
top_hvgs <- getTopHVGs(dec, prop = 0.1)
length(top_hvgs)
```

Save this list of genes as an rds:

```{r}
saveRDS(top_hvgs, file = file.path('visium', 'top_hvgs.rds'))
```

# Plot some of those genes

We can plot these now using the log counts, and explore some different genes:

```{r}
gene <- top_hvgs[20]
plotVisium(spe,
           annotate = gene, 
           point_size = 1, 
           image_ids="hires_bw", 
           assay="logcounts")
```

At this point it might be nice to alter our rownames, 
allowing us to select by gene symbol:

```{r}
rowData(spe)$gene_id <- rownames(spe)
rownames(spe) <- rowData(spe)$symbol
plotVisium(spe,
           annotate = "SPP1", 
           point_size = 1, 
           image_ids="hires_bw", 
           assay="logcounts")
```