---
title: "R Notebook"
output: 05_visium_clustering
editor_options: 
  chunk_output_type: console
---

# Dimensionality reduction algorithms

We have so many features (genes), we can't possibly visualize them all.
We have to reduce the dimensionality of our data.

Let's load our already qc-d SPE.


```{r}
library(SpatialExperiment)
library(ggspavis)
library(scater)
library(scran)
library(pheatmap)

spe <- readRDS(file.path('visium', 'visium_spe_qc.rds'))

top_hvgs <- readRDS(file.path('visium', 'top_hvgs.rds'))

# SVGs defined via nnSVG
#top_svgs <- readRDS(file.path('visium', 'top_svgs.rds'))
```

We're going to use PCA from scater.

```{r}
set.seed(101) # setting a seed makes the following analysis repeatable
spe <- runPCA(spe, subset_row = rownames(top_hvgs))
reducedDimNames(spe)
```

We can also easily use UMAP:

```{r}
set.seed(101) # setting a seed makes the following analysis repeatable
spe <- runUMAP(spe, dimred = "PCA")
reducedDimNames(spe)
plotDimRed(spe, plot_type = "UMAP")
```

At this point, this becomes more art than science...

# Clustering


We're going to use K-nearest neighbors clustering using scran's `buildSNNGraph`.

```{r}

k <- 100 # number of nearest neighbours - higher generally better
# Build a graph to walk on from the UMAP result
g <- buildSNNGraph(spe, k = k, use.dimred = "UMAP")
# Walk this graph, identify group membership
g_walk <- igraph::cluster_walktrap(g)
clus <- g_walk$membership
# See cluster membership counts
table(clus)
# Store cluster labels in the SPE
colLabels(spe) <- factor(clus)
```

Now we can visualise our clusters:

```{r}
plotSpots(spe, annotate = "label", point_size = 1)
```

We can also visualise this back on our UMAP or PCA:

```{r}
plotDimRed(spe, plot_type = "PCA", annotate = "label")
```

Save our SPE with clusters included:

```{r}
saveRDS(spe, file = file.path('visium', 'visium_spe_clust.rds'))
```


# Plot DEGs between clusters.


Finally, let's do some DEG analysis.
It looks like we've got a well-defined tumour area, so let's start there.

```{r}
# Set rownames to symbol rather than gene id
rownames(spe) <- rowData(spe)$symbol

# Test for marker genes, with a dataframe per cluster
# Inspect these dataframes for pvale, FDR, logFC, etc
markers <- findMarkers(spe, test = "binom", direction = "up")
markers
```

# Identify top differences in one of our clusters vs. all other clusters:

```{r}
cluster_of_interest <- markers[[1]]
# Rank here is the minimum rank for the gene across all pairwise comparisons
top_DEGs <- cluster_of_interest[cluster_of_interest$Top <= 10, ]
# Extract effect sizes
logFCs <- getMarkerEffects(top_DEGs)
# Plot heatmap of these changes
pheatmap(logFCs, breaks = seq(-5, 5, length.out = 101))
```

