---
title: "04_1_SVG"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

# This notebook demonstrates how to use nnSVG (optional!)

```{r}
library(SpatialExperiment)
library(ggspavis)
library(nnSVG)

spe <- readRDS(file.path('visium', 'visium_spe_qc.rds'))
```

# Spatially-aware feature selection

We're going to use nnSVG to define spatially-variable genes.

This particular method is quite slow and memory intensive,
So we're going to aggressively subset the data.

```{r}

# Take a random sample of spots
n <- 100
set.seed(101) # set seed to get same random set each time
spe_nnSVG <- spe[, sample(seq_len(n), n)]
```

At this point, we can also do some gene-level filtering, e.g. removing lowly
expressed genes. This function from nnSVG will retain genes with at least ncounts
in pcspots percent of spots.

```{r}
# this function requires "gene_name" to be in rowdata..
# I've set this really high, just because I want it to go faster.
rowData(spe_nnSVG)$gene_name <- rowData(spe_nnSVG)$symbol
spe_nnSVG <- filter_genes(
  spe_nnSVG, filter_genes_ncounts = 20, filter_genes_pcspots = 5
)
```

# Normalization

recalculate log-transformed normalized counts.

We have to run it after doing a gene filter, and also it is required by nnSVG.
```{r}
spe_nnSVG <- logNormCounts(spe_nnSVG)
```

# Spatially variable genes

Be careful running this.
It will eat up loads of memory if you pump n_threads up high.

```{r}
spe_nnSVG <- nnSVG(spe_nnSVG, n_neighbors = 5, n_threads = 2)
```

```{r}
top_svgs <- rowData(spe_nnSVG)[order(rowData(spe_nnSVG)$rank, decreasing = FALSE),]

# let's save the top 300 ranked SVGs
saveRDS(top_svgs[1:300,], file = file.path('visium', 'top_svgs.rds'))
```
